Agile Monitoring of Evolving Web Services: For the Little Man

### 2018-04-07 grouping.tex
cleaning up the introduction.
all is nice up to V

\section{Per-User Performance Monitoring}
\label{sec:user}

It is sometimes the case that the utilization and performance of an API must be understood not only per endpoint, but also by grouping the requests by another criterion. Two such examples are: 
\begin{itemize}
	\item When a service has diferent types of users (e.g. the paying vs. the free-for-evaluation ones, or the in-house vs. those in a different organization) the maintainer must understand the performance for the different groups
	\item When the load mix o the users can vary dramatically and the system response time is a function of the individual user load\footnote{E.g. in GMail some users have a few dozen emails while other have tens of thousands: this difference in user loads will eventually induce a difference in the response times for different users} the maintainer must understand the variation of performance with the users (and implicitly the user load). s
\end{itemize}

@@

After configuring the \tool as shown earlier, the \perspective{Per-User Performance} perspective becomes available to  present the different response times for different users. Figure \ref{fig:tpu} presents a subset of the corresponding view in the \tool. The figure shows that the response times for this endpoint varied considerably for different users with some extreme cases where a user has to wait a full minute until their recommended articles were shown.

>>>>

\section{Identity-Based Grouping of Requests}
\label{sec:grouping}

It is sometimes the case that the utilization and performance of an API must be understood not only per endpoint, but also by grouping the requests by another criterion. Two such examples are: 
\begin{itemize}
	\item When a service has diferent types of users (e.g. the paying vs. the free-for-evaluation ones, or the in-house vs. those in a different organization) the maintainer must understand the performance for the different groups
	\item When the load mix o the users can vary dramatically and the system response time is a function of the individual user load\footnote{E.g. in GMail some users have a few dozen emails while other have tens of thousands: this difference in user loads will eventually induce a difference in the response times for different users} the maintainer must understand the variation of performance with the users (and implicitly the user load).
\end{itemize}

@@

After configuring the \tool as shown earlier, the \perspective{Per-User Endpoint Performance} perspective becomes available to  present the different response times for different users. Figure \ref{fig:tpu} presents a subset of the corresponding view in the \tool. The figure shows that the response times for this endpoint varied considerably for different users with some extreme cases where a user has to wait a full minute until their recommended articles were shown.

### 2018-04-07 outliers.tex
all is nice up to V

  When an API endpoint is called from within a highly interactive application (as it is the case with the case with \epTranslationsColor in the previous section) of particular interest to the API developers are performance {\em outliers}.   Indeed, a translation request that takes three times more than expected can seriously decrease the perceived quality of the application. Thus, identifying, collecting all appropriate data, and diagnosing the root causes of such outliers is especially critical in improving the quality of an application.

  % In the context of the RESTful services support by Flask, this means request serving times that deviate from the average to an unexpected degree.

  For this purpose the \tool tracks for every monitored endpoint a {\em running average} response time value\footnote{\ins{For performance reasons, we assume that the response times for the endpoints are normally distributed. Otherwise, more general density distribution information must be collected in real time.}}. When it detects that a given request is an outlier with respect to this past average running value, it triggers the {\em outlier data collection routine} which stores \ins{extra information} about the current execution environment. A configurable threshold with a default value of $2.5$ times the running average response time is used for this purpose.

  For every detected outlier request, the \tool collects information about the current Python stack trace, CPU load, memory consumption, request parameters, etc. in order to allow the maintainer to investigate the causes of these exceptionally slow response times. In this way it is possible to get \ins{detailed insight into the operation of the application in the extreme cases without unnecessarily burdening it with logging this information for every request}.

  \begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\linewidth]{outlier-annotated}
    \caption{Automatically collected outlier information}
    \label{fig:stack}
  \end{figure}

The bottom panel in \Fref{fig:stack} shows the stack trace. 
In this particular case, it is revealing for the developer to learn that at the time of the stack trace snapshot, the code was in the \code{google\_translator}: indeed, the system uses as back-end multiple translators, and it has been observed that many of the outliers happen while the system is waiting for the Google translator.
%
This information has to be corroborated with the observations that neither the memory nor the processor are overloaded at the moment. Thus this functionality in \code{microsoft\_translator} is really slow in itself, and this is not a result of the machine being overloaded for example. \va{last sentence does not parse}

>>>>

  When an API endpoint is called from within a highly interactive application 
    (as it is the case with \epTranslationsColor in the previous section) 
  of particular interest to the API developers are {\em performance outliers}.   
  Indeed, a translation request that takes three times more than expected can   decrease the perceived quality of the application. 
  Thus, identifying, collecting all appropriate data, and diagnosing the root   causes of such outliers is especially critical in improving the quality of an application.

  For this purpose the \tool tracks for every monitored endpoint a {\em running average} response time value.\footnote{
    For performance reasons, we assume that the response times for the endpoints are normally distributed. 
    Otherwise, more general density distribution information must be collected in real time.}
  When it detects that a given request is an outlier with respect to this past average running value\footnote{A configurable threshold with a default value of $2.5$ times the running average response time is used for this purpose}, it triggers the {\em outlier data collection routine} which stores extra information about the current execution environment. This extra data includes: the current Python stack trace, CPU load, memory consumption, request parameters, etc.

  \begin{figure}[h!]
    \centering
    \includegraphics[width=\linewidth]{outlier-annotated}
    \caption{Automatically collected outlier information for the \epTranslationsColor endpoint}
    \label{fig:stack}
  \end{figure}

\Fref{fig:stack} shows an example of a \perspective{Performance Outlier Detail}  perspective on one of the outliers in the \epTranslationsColor endpoints. In this particular case, the analyst could corroborate several bits of informations:

\begin{itemize}

  \item Neither the memory (1) nor the processor (2) were overloaded at the moment of the request. Thus the slow response is not due to the machine being overloaded.

  \item The stack trace snapshot shows that code was in the \code{google\_translator.py}. Since the system uses as back-end multiple translators it is revealing to see that in this particular case it was waiting for the Google Translator. By manually investigating multiple outliers, the developer discovered that in a large percentage of the cases the Google Translator was indeed the reason for the slow response time.

\end{itemize}

  \niceseparator

  In this way it is possible to get detailed insight into the operation of the application in the extreme cases without unnecessarily burdening it with logging this information for every request. In the discussion section we will come back and discuss some of the limitations of this approach.\ml{memento!}
